---
layout: post
title: "Building a pokemon graph database"
output: github_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
image: '/assets/img/'
description: 'Building a pokemon graph with R and Neo4j'
tags:
- R 
- Neo4j 
categories:
- Pokemon
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What happens when you combine Pokemon with Neo4j?

I'm a huge Pokemon fan. So, when I found about [this awesome post](http://jkunst.com/r/pokemon-visualize-em-all/) from *Joshua Kunst*, I just couldn't wait to throw all that data into Neo4j.

It also happens to be a great way to learn how to build a graph database from scratch. The objective of this exercise is to build a graph database where the nodes are the pokemon and the types, and the relationships are the effectiveness between the pokemon based only on their types.


## Getting the data

First of all, be sure to check Joshua's post to learn how to import all that pokemon data. We will asume that the data is in a data frame called *df*.

Then, we need to get the relationships between types. The easiest thing for acomplishing that is to scrape the table from [pokemondb.net](http://pokemondb.net/type).

```{r, message=FALSE, results='asis'}
library(RNeo4j)
library(rvest)
library(methods)
library(dplyr)

link <- "http://pokemondb.net/type"

link_html <- read_html(link)

types <- link_html %>%
  html_nodes("table") %>%
  .[[1]] %>%
  html_table()

#Give format

names(types)[1] <- "Type"
types$Type <- tolower(types$Type)
names(types)[2:ncol(types)] <- types$Type
types[is.na(types)] <- 1
types[types == ""] <- 1
types[types == "½"] <- 0.5

knitr::kable(types, format = "html")

```

Then we need to separate the types of the pokemon.

```{r, eval=FALSE, echo = TRUE}

df %>% select(id, type =  type_1) -> t1
df %>% select(id, type =  type_2) -> t2

rbind(t1,t2) -> tf

poke_df <- df %>% select(-type_1, -type_2) %>% 
  left_join(tf, by = "id") %>% 
  filter(!is.na(type))

```
We are ready to import to Neo4j, so we need to set the connection.


```{r, eval=FALSE, echo = FALSE}
username <- "Pokemon"
url <- "http://pokemon.sb10.stations.graphenedb.com:24789/db/data/"
password <- "4FUa969pxohWdmBk62qH"

```

Then, we create the pokenodes and the type nodes. We set a relationship for the typing.


```{r, eval=FALSE, echo = TRUE}
#Connect to Graph


graph = startGraph(url = url,
                   username = username,
                   password = password)

#Constraints

addConstraint(graph, "Pokemon", "id")
addConstraint(graph, "Type", "type")


#Create nodes and relationships within the same function

pokenodes <- function(x) {
  pokemon <- getOrCreateNode(graph, "Pokemon", id = x["id"], name = x["pokemon"],
                             height = x["height"], weight = x["weight"],
                             attack = x["attack"], defense = x["defense"],
                             hp = x["hp"], special_attack = x["special_attack"],
                             special_defense = x["special_defense"], speed = x["speed"],
                             url_image = x["url_image"], url_icon = x["url_icon"])
  
  type <- getOrCreateNode(graph, "Type", type = x["type"])
  
  createRel(pokemon,"TYPE",type)
}

#Apply to every row

apply(poke_df[1:nrow(poke_df),],1,pokenodes)

```

We define the desired relationship (effectiveness) using the scraped table


```{r, eval=FALSE, echo = TRUE}


types <- types %>% gather(Type)

names(types)[2] <- "Type_Rel"

effectiveness <- types %>% filter(value != 1)
```

And we are ready to upload the effectiveness, this time using a transaction. Thanks to Nicloe White for this [useful post](https://nicolewhite.github.io/2014/09/30/upload-last-fm-rneo4j-transaction.html)

```{r, eval=FALSE, echo = TRUE}
#Query for creating relationships for the pokenodes

query = "
MERGE (n:Type {type:{type_1}})
MERGE (m:Type {type:{type_2}})
CREATE (n)-[r:EFECTIVENESS]->(m)
SET r.value = {value}
"

#Transactiopn endpoint
t = newTransaction(graph)

for (i in 1:nrow(effectiveness)) {
  type_1 = effectiveness[i, ]$Type
  type_2 = effectiveness[i, ]$Type_Rel
  value = effectiveness[i, ]$value
  
  appendCypher(t, 
               query, 
               type_1 = type_1, 
               type_2 = type_2, 
               value  = value)
}

commit(t)


```

It's time to query our database!!! Let's check all the pokemon that Salamence is double effective:

```{r, echo=FALSE, eval=TRUE}
username <- "Pokemon"
url <- "http://pokemon.sb10.stations.graphenedb.com:24789/db/data/"
password <- "4FUa969pxohWdmBk62qH"

graph = startGraph(     url = url,
                        username = username,
                        password = password)
```

```{r, echo=TRUE, eval=TRUE}
library(visNetwork)

#Query to check for effectiveness for Salamence
final_query <- "
match (n:Pokemon)-[t:TYPE]->(l:Type)-[e:EFECTIVENESS]->(s:Type)<-[j:TYPE]-(z:Pokemon) 
where n.name = 'salamence' 
return n.name as poke1, e.value as value, z.name as poke2, n.url_icon as icon1,
z.url_icon as icon2, n.url_image as image1, z.url_image as image2"

#Execute the query
poke_cypher = cypher(graph, final_query)

#Get data for VisNetwork
poke_cypher <- poke_cypher %>%
  mutate(value = as.numeric(value)) %>%
  group_by(poke1, poke2, image1, image2, icon1, icon2) %>%
  summarise(value = prod(value)) %>%
  ungroup()

#Filter by double effective
poke_sp_eft <- poke_cypher %>%
  filter(value == 2)

#More data for VisNetwork
poke <- unique(c(poke_sp_eft$poke1, poke_sp_eft$poke2))
img  <- unique(c(poke_sp_eft$icon1, poke_sp_eft$icon2))

nodes <- data.frame(id = poke, label = poke, image = img, shape = "image")

edges <- poke_sp_eft %>%
  select(from = poke1, to = poke2)

#The VISUALIZATION
visNetwork(nodes, edges, width = "100%")
```

And that's how you do it! With the RNeo4j it's so easy to set a graph. Maybe in the future it could be expanded in a recommender system or something like that.

Check out a [shiny app](https://jean-arreola.shinyapps.io/Pokemon_Effectiveness/) for the pokemon database! 


